Item Construction Rules Revisited: Learnings from Measurement of Latent Variables with Gold Standard Items

### 2025-07-18 analysis/primary_analyses.qmd
Built site for gh-pages

```{r}
library(foreign, quietly = TRUE, warn.conflicts = FALSE)
library(tidyverse, quietly = TRUE, warn.conflicts = FALSE)
library(semTools, quietly = TRUE, warn.conflicts = FALSE)
library(lavaan, quietly = TRUE, warn.conflicts = FALSE)
library(MBESS, quietly = TRUE, warn.conflicts = FALSE)
library(corx)
```

@@

```{r}
(r_real_poly <- corx(data = dat_poly |>
    select(Alter_realW, Koerpergroesse_realW, Koerpergewicht_realW),
  method = "pearson", stars = c(0.05, 0.01, 0.001)))

r_real_AG6 <- r_real_poly$apa["Alter_realW", "Koerpergroesse_realW"]
r_real_AGe6 <- r_real_poly$apa["Alter_realW", "Koerpergewicht_realW"]
r_real_GGe6 <- r_real_poly$apa["Koerpergroesse_realW", "Koerpergewicht_realW"]

(r_real_dich <- corx(data = dat_dich |>
    select(Alter_realW, Koerpergroesse_realW, Koerpergewicht_realW),
  method = "pearson", stars = c(0.05, 0.01, 0.001)))

r_real_AG2 <- r_real_dich$apa["Alter_realW", "Koerpergroesse_realW"]
r_real_AGe2 <- r_real_dich$apa["Alter_realW", "Koerpergewicht_realW"]
r_real_GGe2 <- r_real_dich$apa["Koerpergroesse_realW", "Koerpergewicht_realW"]
```

@@

```{r}
cor_table <- tribble(
  ~Measure, ~Height_2  , ~Height_6  , ~Weight_2    , ~Weight_6    , ~Age_2       , ~Age_6       ,
  "Height", rel_G2     , rel_G6     , r_latent_GGe2, r_latent_GGe6, r_latent_AG2 , r_latent_AG6 ,
  "Weight", r_real_GGe2, r_real_GGe6, rel_Ge2      , rel_Ge6      , r_latent_AGe2, r_latent_AGe6,
  "Age"   , r_real_AG2 , r_real_AG6 , r_real_AGe2  , r_real_AGe6  , rel_A2       , rel_A6      
)
```

>>>>

```{r}
library(foreign, quietly = TRUE, warn.conflicts = FALSE)
library(tidyverse, quietly = TRUE, warn.conflicts = FALSE)
library(semTools, quietly = TRUE, warn.conflicts = FALSE)
library(lavaan, quietly = TRUE, warn.conflicts = FALSE)
library(MBESS, quietly = TRUE, warn.conflicts = FALSE)
```

@@

```{r}
(r_phys_poly <- standardizedSolution(fit1ges6, type = "std.all") |>
  filter(op == "~~" & 
      lhs %in% c("Alter_realW", "Koerpergroesse_realW", "Koerpergewicht_realW") &
      rhs %in% c("Alter_realW", "Koerpergroesse_realW", "Koerpergewicht_realW") &
      lhs != rhs) |>
  mutate(est_print = papaja::printnum(est.std, gt1 = FALSE, digits = 3)) |>
  mutate(est_print_star = case_when(
    pvalue < 0.05 & pvalue > 0.01 ~ paste0(est_print, "*"),
    pvalue < 0.01 & pvalue > 0.001 ~ paste0(est_print, "**"),
    pvalue < 0.001 ~ paste0(est_print, "***"),
    .default = est_print 
  )))

r_phys_AG6 <- r_phys_poly |> filter(lhs == "Alter_realW" & rhs == "Koerpergroesse_realW") |> pull(est_print_star)
r_phys_AGe6 <- r_phys_poly |> filter(lhs == "Alter_realW" & rhs == "Koerpergewicht_realW") |> pull(est_print_star)
r_phys_GGe6 <- r_phys_poly |> filter(lhs == "Koerpergewicht_realW" & rhs == "Koerpergroesse_realW") |> pull(est_print_star)

(r_phys_dich <- standardizedSolution(fit1ges2, type = "std.all") |>
  filter(op == "~~" & 
      lhs %in% c("Alter_realW", "Koerpergroesse_realW", "Koerpergewicht_realW") &
      rhs %in% c("Alter_realW", "Koerpergroesse_realW", "Koerpergewicht_realW") &
      lhs != rhs) |>
  mutate(est_print = papaja::printnum(est.std, gt1 = FALSE, digits = 3)) |>
  mutate(est_print_star = case_when(
    pvalue < 0.05 & pvalue > 0.01 ~ paste0(est_print, "*"),
    pvalue < 0.01 & pvalue > 0.001 ~ paste0(est_print, "**"),
    pvalue < 0.001 ~ paste0(est_print, "***"),
    .default = est_print 
  )))

r_phys_AG2 <- r_phys_dich |> filter(lhs == "Alter_realW" & rhs == "Koerpergroesse_realW") |> pull(est_print_star)
r_phys_AGe2 <- r_phys_dich |> filter(lhs == "Alter_realW" & rhs == "Koerpergewicht_realW") |> pull(est_print_star)
r_phys_GGe2 <- r_phys_dich |> filter(lhs == "Koerpergewicht_realW" & rhs == "Koerpergroesse_realW") |> pull(est_print_star)
```

@@

```{r}
cor_table <- tribble(
  ~Measure, ~Height_2  , ~Height_6  , ~Weight_2    , ~Weight_6    , ~Age_2       , ~Age_6       ,
  "Height", rel_G2     , rel_G6     , r_latent_GGe2, r_latent_GGe6, r_latent_AG2 , r_latent_AG6 ,
  "Weight", r_phys_GGe2, r_phys_GGe6, rel_Ge2      , rel_Ge6      , r_latent_AGe2, r_latent_AGe6,
  "Age"   , r_phys_AG2 , r_phys_AG6 , r_phys_AGe2  , r_phys_AGe6  , rel_A2       , rel_A6      
)
```

### 2025-07-18 analysis/secondary_analyses.qmd
Built site for gh-pages

---
title: "Secondary analyses"
format:
  html:
    theme:
      - cosmo
      - ../notebooks.scss
    grid:
      body-width: 900px
---

>>>>

---
title: "Secondary analyses (appendices)"
format:
  html:
    theme:
      - cosmo
      - ../notebooks.scss
    grid:
      body-width: 900px
---

::: {.callout-note appearance="simple"}

In this notebook, we record the analyses that are reported in the appendices of the manuscript.

The section [Save results for manuscript](primary_analyses.qmd#sec-saveresults) lists all results objects and tables that are exported to be used in the main manuscript file and can be used to retrace all reported numbers.

:::

### 2025-07-18 manuscript.qmd
Built site for gh-pages

[^notewording]: Compared to the pilot test, one height and one weight item were slightly reworded to simply the wording and reduce similarities with other items in the scale.

@@

For body height, we obtained low item loadings on the latent variable but high regression weights with the gold-standard item of physical body height. In particular, the third item ("Chairs and tables are usually too low for me.") exhibited the highest loading on the latent variable for both response formats. In contrast, item 11 ("I have to stand at the front of group photos so that I can be seen clearly.") and item 12 ("When I hug other people in greeting, I have to bend downwards.") were best predicted by the self-reported physical body height for both response formats.[^loadingsgreaterone] This pattern does not convincingly confirm the existence of a psychological latent height variable after controlling for the self-reported physical measure and gender.

>>>>

[^notewording]: Compared to the pilot test, one height and one weight item were slightly reworded to simplify the wording and reduce similarities with other items in the scale.

@@

For body height, we obtained low item loadings on the latent variable but high regression weights with the gold-standard item of physical body height. In particular, the third item ("Chairs and tables are usually too low for me.") exhibited the highest loading on the latent variable for both response formats. In contrast, item 11 ("I have to stand at the front of group photos so that I can be seen clearly.") and item 12 ("When I hug other people in greeting, I have to bend downwards.") were best predicted by the self-reported physical body height for both response formats.[^loadingsgreaterone] This pattern does not convincingly confirm the existence of a psychological latent height variable after controlling for the self-reported physical item and gender.
